---
title: "Urban Big Data Analytics - Lecture 10 Group Session: Advanced Modeling"
author: "Andy Hong"
date: "July 31, 2019"
output:
  html_document:
    css: my_styles.css
  pdf_document: default
  word_document: default
---

## Prerequisites

Please locate your `vsp_bigdata` folder under "My Documents" and navigate to `group-session`. Create `10-lecture` folder under the `group-session` folder.

For this group session, we will use the 2011 Gapminder database. <br>
Please download this CSV file and save it under the group session folder: <a href="https://urbanbigdata.github.io/10-lecture-advanced_modeling/group-session/gapminder_data_2011.csv" download> Gapminder data 2011 </a><br>

<br><br>


## Instruction

### 1. Synopsis
The purpose of this group session is to understand the concept of **logistic regression**. <br> 

First, we will build a logistic regression model using Exploratory. Then, we will build a logistic regression model using R Studio.
<br>

### 2. Part one - Building a logistical model in Exploratory

Let's load the gapminder 2011 data. <br>
Each country has a corruption perception index ranging from 0 to 100, and based on this index, we assigned 1 for countries that have the index greater than 50; and 0 for countries that have the index less than 50. The corruption variable is an arbitrary variable only for a demonstration purpose.<br>

Go to `Analytics` and choose `Logistic Regression Analysis` for the type. <br>
Your target variable is `corruption`, and your predictor variable(s) are: `income`, `population`, and `democracy`.<br>

Mathematically, what we are trying to estimate is the following function.<br>

$logit(p) = log(\frac{p}{1-p}) = \beta_{0} + \beta_{1}x_{1} + \beta_{2}x_{2} + \beta_{3}x_{3}$

where $log(\frac{p}{1-p})$ is the log of odds ratio being in corruption; $x_{1}$is income; $x_{2}$is population; and $x_{3}$is democracy. <br> 

When you click run, you will see the coefficient plot of the logit model results.<br>
Take a look at the result of democracy. <b>The odds ratio of 0.79 means that for 1-unit increase in democracy, we expect to see about 21% decrease in the odds of being in corruption.</b>
<br><br>

### 3. Part two - Building a logistical model in R Studio

We are going to join the gapminder data with the corruption data. Let's load the libraries first.

```{r warning=FALSE, message=FALSE}

# Set CRAN repository source
options(repos="https://cran.rstudio.com")

# Install packages
# install.packages("dplyr")
# install.packages("magrittr")

# Load packages
library(dplyr)
library(magrittr)

```

Let's load and examine the gapminder data. 

```{r}
gapminder = read.csv("/Users/andyhong/Documents/vsp_bigdata/group-session/10-lecture/gapminder_data_2011.csv")

head(gapminder)
```

We are going to build the basic and the full models. We will use the `glm` function, which is a generalized linear model function. We will use the binomial family to implement the logistic function.


```{r}

model1 = glm(data = gapminder, corruption ~ income, family = binomial)
model2 = glm(data = gapminder, corruption ~ income + population + democracy, family = binomial)

```

Now we can see the results of the models using the `summary` function.

```{r}
# Show the original results
summary(model1)
summary(model2)
```

For interpretation, we can exponentiate the results to get the odds ratio.
```{r}
# Show the exponentiated results
exp(cbind(Odds=coef(model1), confint(model1)))
exp(cbind(Odds=coef(model2), confint(model2)))
```

Now, we can also generate the coefficient plot using the `plot_summs` function from the `jtools` package. To get the exponentiated results, we will use the option `exp=TRUE`.

```{r}

# Load the jtools package
library(jtools)

# Show coefficient plot
plot_summs(model1, model2, exp=TRUE)

```


## Send your report with the logistic model results and the interpretation

1. Use either Exploratory or R Studio to develop the best fit logistic model to estimate corruption 

2. Create a short report that includes the model results and interpret your results.

3. Send your report to the course email (urbanbigdata2019@gmail.com). 
  + <font color="red"><b>[IMPORTANT]</b></font> Please use the following email title format: <br>
 `VSP BigData [lecture number] - [group number] - [presenter name]` <br>
 ex), `VSP BigData Lecture 11 - Group 1 - Bill Gates` <br>
 
 
<br><br><br>
 

 




